\documentclass[12pt]{ucsddissertation}
% mathptmx is a Times Roman look-alike (don't use the times package)
% It isn't clear if Times is required. The OGS manual lists several
% "standard fonts" but never says they need to be used.
%\usepackage{mathptmx}

\usepackage{fontspec}
\usepackage{gfsdidot}          % makes math stuff look neat, but also overrides text stuff
\setmainfont{TeX Gyre Pagella} % forces back text stuff to look neat too
\setmonofont{DejaVuSansMono}   % monospace stuff should look neat too
%\renewcommand{\familydefault}{\sfdefault}
%\usepackage[scaled=1]{helvet}

\usepackage[NoDate]{currvita}
\usepackage{array}
%\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{ragged2e}
\usepackage{microtype}

\usepackage[numbers]{natbib}
%\usepackage[style=ieee,natbib=true]{biblatex}

\usepackage{graphicx}

% Removed for preliminary appointment
% \usepackage{showframe}

% Valentin
\usepackage[dvipsnames]{xcolor} % must be loaded prior to tikz, pgfplots, etc.
\let\iint\relax
\let\iiint\relax
\let\iiiint\relax
\let\idotsint\relax
\usepackage{amsmath} % loaded by mathtools?
\usepackage{amssymb}
\usepackage[english]{babel}
\usepackage{calc}
\usepackage{caption}
\usepackage{color}
\usepackage{DejaVuSansMono}
\usepackage{dsfont}
\let\textlozenge\relax
\usepackage[inline]{enumitem}
\usepackage{environ}
\usepackage{etoolbox}
\usepackage{fontawesome}
\usepackage{forest}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{ltablex} \keepXColumns % Made me sad for appendix tables, where was it used?
\usepackage{makecell}
\usepackage{mathpartir}
%\usepackage{mathtools} % \coloneqq
\usepackage{mdframed}
\usepackage{minted}
\usepackage{multicol}
\usepackage[parfill]{parskip}
\usepackage{pgfplots}
\usepackage{pifont} % \ding
\usepackage{soul}
\usepackage{syntax}
\usepackage{tabularx}
\usepackage[skins,theorems]{tcolorbox}
\usepackage{tikz}
\usepackage{tikzpeople}
\usepackage{tikz-qtree}
\usepackage{xpatch}
\usepackage[breaklinks=true,pdfborder={0 0 0}]{hyperref}

% gfsdidot makes ∀ and ∃ super ugly, this reverts them to something fine
\DeclareSymbolFont{CMsymbols}{OMS}{cmsy}{b}{n}
\SetSymbolFont{CMsymbols}{bold}{OMS}{cmsy}{ub}{n}
\DeclareMathSymbol{\forall}{\mathord}{CMsymbols}{"38}
\DeclareMathSymbol{\exists}{\mathord}{CMsymbols}{"39}

\definecolor{color01}{HTML}{E6194B}
\definecolor{color02}{HTML}{3CB44B}
\definecolor{color03}{HTML}{FFE119}
\definecolor{color04}{HTML}{4363D8}
\definecolor{color05}{HTML}{F58231}
\definecolor{color06}{HTML}{911EB4}
\definecolor{color07}{HTML}{46F0F0}
\definecolor{color08}{HTML}{F032E6}
\definecolor{color09}{HTML}{BCF60C}
\definecolor{color10}{HTML}{FABEBE}
\definecolor{color11}{HTML}{008080}
\definecolor{color12}{HTML}{E6BEFF}
\definecolor{color13}{HTML}{9A6324}
\definecolor{color14}{HTML}{FFFAC8}
\definecolor{color15}{HTML}{800000}
\definecolor{color16}{HTML}{AAFFC3}
\definecolor{color17}{HTML}{808000}
\definecolor{color18}{HTML}{FFD8B1}
\definecolor{color19}{HTML}{000075}
\definecolor{color20}{HTML}{808080}
\definecolor{color21}{HTML}{FFFFFF}
\definecolor{color22}{HTML}{000000}

\def\mathunderline#1#2{\color{#1}\underline{{\color{black}#2}}\color{black}}

\forestset{
  default preamble={
    for tree={
      align=center,
      draw=black,
      edge={
        ultra thick,
      },
      edge path={
        \noexpand\path[\forestoption{edge}]
        (!u.parent anchor)
        -- +(0,-10pt)
        -| (.child anchor)
        \forestoption{edge label};
      },
      font=\bfseries,
      % inner sep=10pt,
      l sep=30pt,
      % line width=2pt,
      minimum height=20pt,
      minimum width=30pt,
      parent anchor=south,
      rectangle,
      ultra thick,
    }
  }
}

\newcolumntype{Y}{>{\centering\arraybackslash}X}

% defines safecoqinline, a command like coqinline but works in tabular
% environments
\makeatletter
\def\safecoqinline#1{%
\ifx\@footnotetext\TX@trial@ftn
\detokenize{#1}%
\else
\coqinline{#1}%
\fi}
\makeatother

\captionsetup{
  justification=centering,
}

\usetikzlibrary{arrows, calc, fit, positioning, shapes, tikzmark}

\tikzset{
bicolor/.style 2 args={
  dashed,dash pattern=on 4pt off 4pt,#1,
  postaction={draw,dashed,dash pattern=on 4pt off 4pt,#2,dash phase=4pt}
  },
}

\tikzstyle{Matching}=[
black,
dashed,
line width=3pt,
]

\tikzstyle{RoundedDottedPath}=[
densely dotted,
color05,
line width=3pt,
rounded corners=10pt,
]

\tikzstyle{RoundedRectangle}=[
dashed,
draw,
line width=3pt,
rounded corners=10pt,
]

\tikzstyle{NodeLabel}=[
black,
circle,
draw,
fill=white,
font=\bfseries,
inner sep=1pt,
ultra thick,
]

\graphicspath{ {./images/} }

\AtBeginDocument{%
	\settowidth\cvlabelwidth{\cvlabelfont 0000--0000}%
}

% Fix a mdframed bug where skipbelow is ignored
\makeatletter
\xpatchcmd{\endmdframed}
  {\aftergroup\endmdf@trivlist\color@endgroup}
  {\endmdf@trivlist\color@endgroup\@doendpe}
  {}{}
\makeatother

\surroundwithmdframed[
backgroundcolor=white,
linecolor=white,
skipabove=1em,
skipbelow=-0.5em,
leftmargin=10pt,
innertopmargin=1pt,
innerbottommargin=0pt
]{minted}

\setminted{
  % DO NOT use bgcolor, use mdframed instead
  %bgcolor=white,
  linenos=true,
  style=xcode,
  breaklines=true,
  encoding=utf8,
  fontsize=\small,
  baselinestretch=1,
}

% need bgcolor for the display to be nice in rules
\newmintinline{coq}{bgcolor=white}

% This works for some styles but not all, instead I hijack the tokens directly
% in background.tex

% \AtBeginEnvironment{minted}{%
%   \renewcommand{\fcolorbox}[4][]{#4}}
% \def\dontdofcolorbox{\renewcommand\fcolorbox[4][]{##4}}

% These do not seem to work...
% \xpatchcmd{\coqinline}{\minted@fvset}{\minted@fvset\dontdofcolorbox}{}{}
% \xpatchcmd{\mintinline}{\minted@fvset}{\minted@fvset\dontdofcolorbox}{}{}

\makeatletter
% \AtBeginEnvironment{minted}{\dontdofcolorbox}
% \def\dontdofcolorbox{\renewcommand\fcolorbox[4][]{##4}}
% \xpatchcmd{\inputminted}{\minted@fvset}{\minted@fvset\dontdofcolorbox}{}{}
% \xpatchcmd{\mintinline}{\minted@fvset}{\minted@fvset\dontdofcolorbox}{}{}

% \expandafter\def\csname PYGmonokai@tok@err\endcsname{
%   \def\PYGmonokai@tc##1{\textcolor[rgb]{0.59,0.88,0.31}{##1}}
%   \def\PYGmonokai@bc##1{\setlength{\fboxsep}{0pt}\colorbox[rgb]{0.88,0.00,0.06}{\strut ##1}}
% }
\makeatother

\makeatletter
\pgfdeclareshape{document}{
\inheritsavedanchors[from=rectangle] % this is nearly a rectangle
\inheritanchorborder[from=rectangle]
\inheritanchor[from=rectangle]{center}
\inheritanchor[from=rectangle]{north}
\inheritanchor[from=rectangle]{south}
\inheritanchor[from=rectangle]{west}
\inheritanchor[from=rectangle]{east}
% ... and possibly more
\backgroundpath{% this is new
% store lower right in xa/ya and upper right in xb/yb
\southwest \pgf@xa=\pgf@x \pgf@ya=\pgf@y
\northeast \pgf@xb=\pgf@x \pgf@yb=\pgf@y
% compute corner of ‘‘flipped page’’
\pgf@xc=\pgf@xb \advance\pgf@xc by-10pt % this should be a parameter
\pgf@yc=\pgf@yb \advance\pgf@yc by-10pt
% construct main path
\pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@ya}}
\pgfpathlineto{\pgfpoint{\pgf@xa}{\pgf@yb}}
\pgfpathlineto{\pgfpoint{\pgf@xc}{\pgf@yb}}
\pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yc}}
\pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@ya}}
\pgfpathclose
% add little corner
\pgfpathmoveto{\pgfpoint{\pgf@xc}{\pgf@yb}}
\pgfpathlineto{\pgfpoint{\pgf@xc}{\pgf@yc}}
\pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yc}}
\pgfpathlineto{\pgfpoint{\pgf@xc}{\pgf@yc}}
}
}
\makeatother

% OGS recommends increasing the margins slightly.
\increasemargins{.1in}

% These are just for testing/examples, delete them
%\usepackage{trace}
%\usepackage{showframe} % This package was just to see page margins
%\usepackage[english]{babel}
%\usepackage{blindtext}
\overfullrule5pt
% ---

% Required information
\title{Front-end tooling for building and maintaining functional programs}
\author{Valentin Robert}
\degree{Computer Science}{Doctor of Philosophy}
% Each member of the committee should be listed as Professor Foo Bar.
% If Professor is not the correct title for one, then titles should be
% omitted entirely.
\chair{Professor Sorin Lerner}
%\cochair{Professor Gamma Delta} % Optional
% Your committee members (other than the chairs) must be in alphabetical order
\committee{Professor William Griswold}
\committee{Professor James Hollan}
\committee{Professor Ranjit Jhala}
\committee{Professor Todd Millstein}
\degreeyear{2018}

\input{macros}
\input{chick-rules}

\setlength\parindent{0.5in}
%\titleformat\paragraph{\sectioningformatting\normalsize\bfseries}{\theparagraph}{1em}{}

% Start the document
\begin{document}
% Begin with frontmatter and so forth
\frontmatter
\maketitle
\makecopyright
\makesignature
% Optional
% \begin{dedication}
% \setsinglespacing
% \raggedright % It would be better to use \RaggedRight from ragged2e
% \parindent0pt\parskip\baselineskip
% In recognition of reading this manual before beginning to format the
% doctoral dissertation or master's thesis; for following the
% instructions written herein; for consulting with OGS Academic Affairs
% Advisers; and for not relying on other completed manuscripts, this
% manual is dedicated to all graduate students about to complete the
% doctoral dissertation or master's thesis.

% In recognition that this is my one chance to use whichever
% justification, spacing, writing style, text size, and/or textfont that
% I want to while still keeping my headings and margins consistent.
% \end{dedication}
% % Optional
% \begin{epigraph}
% \vskip0pt plus.5fil
% \setsinglespacing
% {\flushright
% True ease in writing comes from art, not chance,\\
% As those move easiest who have learn'd to dance.\\
% 'T is not enough to no harshness gives offence,---\\
% The sound must seem an echo to the sense.

% \vskip\baselineskip
% \textit{Alexander Pope}\par}
% \vfil
% \begin{center}
% You write with ease to show your breeding,\\
% But easy writing's curst hard reading.

% \vskip\baselineskip
% \textit{Richard Brinsley Sheridan}
% \end{center}
% \vfil
% \noindent Writing, at its best, is a lonely life. Organizations for
% writers palliate the writer's loneliness, but I doubt if they improve
% his writing. He grows in public stature as he sheds his loneliness and
% often his work deteriorates. For he does his work alone and if he is a
% good enough writer he must face eternity, or the lack of it, each day.

% \vskip\baselineskip
% \hskip0pt plus1fil\textit{Ernest Hemingway}\hskip0pt plus4fil\null

% \vfil
% \end{epigraph}

% Next comes the table of contents, list of figures, list of tables,
% etc. If you have code listings, you can use \listoflistings (or
% \lstlistoflistings) to have it be produced here as well. Same with
% \listofalgorithms.
\tableofcontents
\listoffigures
\listoftables

% Preface
% \begin{preface}
% \end{preface}

% Your fancy acks here. Keep in mind you need to ack each paper you
% use. See the examples here. In addition, each chapter ack needs to
% be repeated at the end of the relevant chapter.
\begin{acknowledgements}
I would like to thank Professor Sorin Lerner for his support as my advisor and
the chair of my committee.

Committing oneself to more than half a decade to a mentally-challenging,
ill-paying, self-motivated job seems like an awful idea.  Yet, it is the ordeal
that most graduate students choose, with little to no foresight as to where the
journey will lead them.  I would like to dedicate the next paragraphs to all
past, current, and future graduate students.

Throughout my six years as a graduate student, I have only been able to visit my
family for a total of three months, both my grandparents passed away after I was
unable to attend a Christmas with them, nor was I able to attend their
respective funeral.  I have needed to move twice, I have needed to TA seven
quarters, I have developed repetitive stress injuries in both my arms, and I
have loaned over \$10.000 to be able to bring this work to completion.

It was not without its worth though.

I have made a solid group of friends without which the journey would have been
too much to bear.  Zachary Tatlock brought me up to speed and offered me several
wonderful opportunities over the years.  Alexander Bakst was a great roommate,
and it was always great to go for a bike ride to go work the night at one of the
local coffee shops.  Neha Chachra was a great neighbour, and helped me work
through a lot of mental issues.  Marcela Mendoza also provided a lot of mental
support, personal growth, and both helped focus on working hard for several
years, and helped on rejuvenating from the worst grinds by being a great travel
companion.  Many recent new friends have made the final years of the journey
wonderful, with lots of music and enjoyable hangouts.  In particular, my partner
Andrea Frank provided a wonderful emotional support over the past year, and with
her roommate, John Renner, have provided the hosting without which I would not
have been financially able to complete my studies.

I would like to thank the rest of the UCSD Programming Systems group, who have
provided a great environment for research.  In particular, Eric Seidel has
always been tremendous at listening to my issues and offering the perfect,
appropriate technical pointers to resolve them.  My work would often have been
weeks behind without his help.

I would also like to thank the Gallium team of Inria, and in particular my
former mentor, Xavier Leroy, for his great mentorship and the confidence he
inspired in me.  Gabriel Scherer, Jonathan Protzenko, and Nicolas Pouillard,
among many others, have also been excellent lunch and coffee break companions,
who opened my eyes to many of the great concepts of the programming language
community.  Working with all of them was a daily delight.

Andrew Kennedy, who accepted me as an intern at Microsoft Research, gave me the
opportunity to work on a project where I got to improve my understanding of
\Coq{} at a tremendous speed, and introduced me to a project that I had not
believed possible.

I would like to thank the employees of Galois, Inc., where I gave a presentation
of my work earlier this year that was very well received, and followed by great
feedback.  In particular, David Christiansen pointed me to relevant literature
that made part of the work much easier.  It is an immense pleasure to have been
offered a position to work alongside the people I met that day.

Finally, I must thank my mother for her love and support.  This journey must
have been quite the ordeal for her too, and it was hard to convey the reasons
why I still had so much left to do before I was done.  Yet, she was
understanding, and always came through at the most difficult moments, no
questions asked.  Such care is commendable.

Chapter~\ref{peacoq}, in part, has been published in a series of blog
articles. The dissertation author was the primary investigator and author of
this material.

Chapter~\ref{chick}, in part, is currently being prepared for submission for
publication of the material.  The dissertation author was the primary
investigator and author of this material.

% I would like to acknowledge Professor Eta Theta for his support as the
% chair of my committee. Through multiple drafts and many long nights,
% his guidance has proved to be invaluable.

% I would also like to acknowledge the ``Smith Clan'' of lab~28, without
% whom my research would have no doubt taken fives times as long. It is
% their support that helped me in an immeasurable way.

% Chapter 2, in full, is a reprint of the material as it appears in
% Numerical Grid Generational in Computational Fluid Mechanics~2009.
% Smith, Laura; Smith, Jane~D., Pineridge Press,~2009. The dissertation
% author was the primary investigator and author of this paper.

% Chapter 3, in part, has been submitted for publication of the material
% as it may appear in Education Mechanics,~2009, Smith, Laura; Smith,
% Jane~D., Trailor Press,~2009. The dissertation author was the primary
% investigator and author of this paper.

% Chapter 5, in part is currently being prepared for submission for
% publication of the material. Smith, Laura; Smith, Jane~D\@. The
% dissertation author was the primary investigator and author of this
% material.

\end{acknowledgements}

% Stupid vita goes next
\begin{vita}
\noindent
\begin{cv}{}
\begin{cvlist}{}
\item[2012 - 2018] Doctorate of Philosophy, Computer Science, University of California San Diego
\item[2008 - 2012] Diploma of Engineering (Master's degree), ENSEIRB-MATMECA
\item[2006 - 2008] Classe pr\'eparatoire, Lyc\'ee Louis Barthou
\item[2001 - 2006] Baccalaur\'eat, Science specialization, Lyc\'ee Saint-Cricq
% \item[1996] Bachelor of Arts, University of California, Berkeley
% \item[1996--2000] U.S. Marines
% \item[2000--2002] Teaching Assistant, Department of Mechanical
% Engineering\\University of California, San Diego
% \item[2002--2006] Research Assistant, University of California, San
% Diego
% \item[2010] Doctor of Philosophy, University of California, San Diego
\end{cvlist}
\end{cv}

% This puts in the PUBLICATIONS header. Note that it appears inside
% the vita environment. It is optional.
\publications

\noindent Robert, V. and Leroy, X., 2012, December. A formally-verified alias
analysis. In International Conference on Certified Programs and Proofs
(pp. 11-26). Springer, Berlin, Heidelberg.

\noindent Ricketts, D., Robert, V., Jang, D., Tatlock, Z. and Lerner, S., 2014,
June. Automating formal proofs for reactive systems. In ACM SIGPLAN Notices
(Vol. 49, No. 6, pp. 452-462). ACM.

% This puts in the FIELDS OF STUDY. Also inside vita and also
% optional.
% \fieldsofstudy
% \noindent Major Field: Engineering (Specialization or Focused Studies)
% \vskip\baselineskip
% Studies in Applied Mathematics\par
% Professors Alpha Beta and Gamma Delta
% \vskip\baselineskip
% Studies in Mechanices\par
% Professors Epsilon Zeta and Eta Theta
% \vskip\baselineskip
% Studies in Electromagnetism\par
% Professors Iota Kappa and Lambda Mu
\end{vita}

% Put your maximum 350 word abstract here.
\begin{dissertationabstract}

Dependently-typed functional languages are increasingly popular, but due to the
complexity of their type systems, there is still a lot of friction in the user
experience, both for beginners who try to learn the concepts, and expert users
who must write and maintain complex code bases.  We explore ways to alleviate
those burdens by providing novel front-end tooling for this class of languages.

In order to help beginners, we explore new visualizations and automation
techniques, focusing on three pain points we identified in the learning process.
We evaluate, via a longitudinal user study and an A-B study, their effectiveness
in terms of learning to use those languages, enjoyment of the learning process,
and productivity on solving beginner-level exercises.

In order to help experts, we prototype a tool that helps in the refactoring of
programs, partially eliminating the tedium of propagating changes throughout
large code bases.  We evaluate this tool over a choice of refactoring
benchmarks, and show how it can be extended to support other languages than the
toy language we designed it for.  We evaluate one instance of this extension
technique to repair programs in the \OCaml{} language, a functional language
without dependent types, but with more features than our tool was built for.

% The Abstract begins here. The abstract is limited to 350 words for a
% doctoral dissertation. It should consist of a short statement of the
% problem, a brief explanation of the methods and procedures employed in
% generating the data, and a condensed summary of the findings of the
% study. The abstract may continue onto a second page if necessary. The
% text of the abstract must be double spaced.
\end{dissertationabstract}

% This is where the main body of your dissertation goes!
\mainmatter

% Optional Introduction
\begin{dissertationintroduction}

\input{introduction}

% This optional section is barely described in the OGS manual other than
% saying it is optional and that it appears in the table of contents
% between the Abstract and the first chapter.

% No formatting guidelines appear so presumably, it should be formatted
% like an ordinary chapter. It should appear after the
% \verb!\mainmatter! macro because it should start on page~1.
\end{dissertationintroduction}

% Too much
%\input{background}
%\input{peacoq}
\input{chick}
%\input{related-work}

%\chapter{Takeaways}

\appendix

%\include{appendix}

% Stuff at the end of the dissertation goes in the back matter
\backmatter
\bibliographystyle{plainnat} % Or whatever style you want like plainnat
\bibliography{bibliography}

\end{document}
