\section{Background}

A proof assistant is typically composed of several interacting pieces:

\begin{itemize}

  \item a \emph{programming language}, allowing the user can define programs
that they wish to execute and/or reason about formally,

  \item a \emph{specification language}, allowing the user to define properties
of programs, or general theorems, that they wish to prove,

  \item a \emph{proof language}, allowing the user to build those proofs.

\end{itemize}

Note that these languages need not be different from one another, and several
languages can help fulfill one of those purposes.  For instance, in the Coq
proof assistant, the programming language and the specification language are the
same language, called \Gallina{}, and the proof language can be either
\Gallina{} itself, or a proof-building scripting language called \Ltac{}.

Learning to use such a proof assistant therefore requires familiarizing oneself
with not only a programming language with a complex type system, but also the
logical foundation of formal proofs, and the idiosyncrasies of the proof
environment of choice.

For the Coq proof assistant, novice users are generally taught to use \Ltac{} to
build proofs, since manually building \Gallina{} terms requires more expertise
and is usually tedious.  An \Ltac{} script is a sequence of commands (usually
called \define{tactics}) directing the proof assistant as to what logical rules
to use in order to progress in building the proof term.  These tactics include:

\begin{itemize}

  \item proof-solving tactics, which attempt to terminate a sub-proof by
constructing the complete proof term, therefore finishing this part of the
proof,

  \item case-splitting tactics, which break a proof obligation into several
sub-obligations, by using some rule of the formal logic with multiple
antecedents,

  \item bookkeeping tactics, which modify the context or the goal of the current
proof obligation, either by adding/removing/reordering/rewriting in hypotheses,
or by performing modifications in the goal.

\end{itemize}

A typical proof can therefore be thought of as a tree, branching on
case-splitting tactics, extending on bookkeeping tactics, and with proof-solving
tactics as leaves.  The proving process is rarely linear: many proofs will
require using concepts such as \emph{case analysis} and \emph{induction} in
order to break down a complex goal into specialized sub-goals that can be solved
separately.  In those sub-cases, one will often need to perform applications and
rewriting using existing theorems.

The \define{application} of a theorem can be performed on either a hypothesis or
a goal.  A theorem or hypothesis can be applied to a hypothesis if the
hypothesis matches one of the antecedents of the theorem.  It then yields the
conclusion of the theorem as a hypothesis, but also adds new obligations for all
the other antecedents. That is, given the context:

\begin{minted}{coq}
A, B, C : Prop
H1 : A → B → C
H2 : B
==================
C
\end{minted}

applying hypothesis \mintinline{coq}{H1} to hypothesis \mintinline{coq}{H2}
yields the following two obligations:

\begin{minted}{coq}
A, B, C : Prop
H1 : A → B → C
H2 : C
==================   (* 1. Hypothesis H2 became the conclusion of H1. *)
C

A, B, C : Prop
H1 : A → B → C
H2 : B
==================   (* 2. In the original context, one must prove A. *)
A
\end{minted}

Conversely, a theorem or hypothesis can be applied to the goal of the current
obligation if its conclusion matches the goal.  It yields an obligation for each
antecedent.  For instance, applying hypothesis \mintinline{coq}{H1} from the
original context to the goal would yield the following two obligations:

\begin{minted}{coq}
A, B, C : Prop
H1 : A → B → C
H2 : B
==================   (* 1. One must prove the first antecedent A. *)
A

A, B, C : Prop
H1 : A → B → C
H2 : B
==================   (* 2. One must prove the second antecedent B. *)
B
\end{minted}

% TODO

We identified three challenges to beginners in the learning process:

\begin{enumerate}

  \item conceptualizing and keeping track of the \emph{proof tree structure}
while building proofs,

  \item identifying the effects of an \Ltac{} tactic on the proof context,

  \item identifying tactics

\end{enumerate}
